Between Cell Variance

### 2020-05-18 Chap3_MLH1var.rmd
corrected permutation and replicate number

#library(tidyverse)
library(knitr)
library(ggplot2)
#library(pwr)
library(purrr) #for replicate
library(dplyr)
library(plyr)
library(lattice)

library(raster)
library(lme4)
library(nlme)
library(RLRsim)
#library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

@@

Answering these will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution

@@

# Q12 Varince Plots

@@

The figure above nominates MOLF should be removed from the full dataset. I think all cells can be included for the sex specific data.

The female LEW strains have a larger distribution of variance. (or it could just be the 2 mice). Some of the low cell number mouse.  (note on the dot size, small dots could be mice with few high Q cells. Med size cells have a higher number of cells.)

@@

#strain = fixed, 
varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

@@

**M4-M5 male**
```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

@@

**M2**
```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain *Ncells, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

@@

**M3**
```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

@@

**M4-M5 female**
```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

@@

**M4-M5 male**
```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

@@

sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)
round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)

@@

I would suggest the following approach:
1. Consider only high-quality cells.
2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.

@@

4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
Compute the among-strain variance in the MLH1 count variance (separately for each sex) in your original data.
Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).
-->

@@

Calculate the category level between cell variance (and cV).

@@

@@

@@

Made a table with the cell-pooled calculations of var and cV.
Next step is to i) record the data structure ii) then permute within sex, then calculate the cell-pooled values.

```{r permutations, echo=FALSE, eval=FALSE}

Q12.permute.DF <- Q12.MLH1_data %>% filter()

remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]

#Data structure
len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

@@

#these are the bones of the structure -- for permuting a new DF -- then calculating

#this DF is kinda hard, I'm not sure how to make a full randomized DF

#replicate(n = 3, rnorm(5, 0, 1), simplify = FALSE )

nrepeat = 50

permute.var <- data.frame( "wsb.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.WSB.female, replace = FALSE) ) ),

                            "wsb.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
  "lew.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.LEW.female, replace = FALSE) ) ),

  "lew.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.LEW.male, replace = FALSE) ) ),

  "g.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.G.female, replace = FALSE) ) ), 
  "g.m.var" =  replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.G.male, replace = FALSE) ) ),

  "pwd.f.var" = replicate(n=nrepeat, var( sample( Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
  "pwd.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.PWD.male, replace = FALSE))),

  "kaz.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.KAZ.female,replace = FALSE) ) ),

  "kaz.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.KAZ.male, replace = FALSE)) ),

  "skive.f.var" = replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),

  "skive.m.var" =  replicate(n=nrepeat,var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),
  "msm.f.var" = replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm.m.var" =  replicate(n=nrepeat, var(sample(Q12.permute.DF$nMLH1.foci[Q12.permute.DF$sex == "male"], len.MSM.male, replace = FALSE) )   )  )

#now that these is a dataframe of new/permuted variance (specific to each sex) ... compare them to the origincal values?
#my prediction is that there would be many significant differences from the true values

#to visualize, I can plot the variances .. I guess with a histogram

#now automate making these plots -- or calculate the propotion of permutations which are below the observed

pp <- ggplot(permute.var, aes(x = wsb.f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )

#randomized data frame, sample mouse means from all P.leucopus mice

#calculate pvalue as proportion of differences greater than observed difference
p.val_permut_mean.lab.wild <- length(permute.dif.meanOmeans$permuted.diff[permute.dif.meanOmeans$permuted.diff >= obs.dif.mean.lab.wild]) / permut.x
var.pval <- paste( "p-value ", p.val_permut_mean.lab.wild)

#visualize the permuted distribution
permute_hist_means  <-  ggplot(permute.dif.meanOmeans,aes(x=permuted.diff))+
  geom_histogram()+geom_vline(xintercept=obs.dif.mean.lab.wild,color='red') +
  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of mouse means, wild to lab")

@@

#Same permutations, at the cell level
Pl.cells <- MLH1_data.hq[MLH1_data.hq$species == "Pleuc",]

obs.status.cell.diff <- mean(Pl.cells$nMLH1.foci[Pl.cells$status == "wild"])-
    mean(Pl.cells$nMLH1.foci[Pl.cells$status == "lab"])

permute.dif.cell.status  <- as.data.frame( replicate(permut.x, 
                                  (  mean(sample(Pl.cells$nMLH1.foci,  243, replace = FALSE) ) - 
                                   mean(sample(Pl.cells$nMLH1.foci, 221, replace = FALSE) ) )  
) )

colnames(permute.dif.cell.status) <- "permuted.diff"
p.val_permut_cell.diff.lab.wild <- length(permute.dif.cell.status$permuted.diff[permute.dif.cell.status$permuted.diff >= obs.status.cell.diff]) / permut.x
var.pval <- paste( "p-value ", p.val_permut_cell.diff.lab.wild)

permute_hist_means.cell_level  <-  ggplot(permute.dif.cell.status, aes(x=permuted.diff))+
  geom_histogram()+geom_vline(xintercept=obs.status.cell.diff,color='red') +
  annotate("text", x= 0, y = 75, label =  var.pval)+ggtitle("permuted difference of cell level means, wild to lab")

>>>>

#library(tidyverse)
library(knitr)
library(ggplot2)
#library(pwr)
library(purrr) #for replicate
library(dplyr)
library(plyr)
library(lattice)
library(tidyr)
library(raster)
library(lme4)
library(nlme)
library(RLRsim)
#library(ggpubr)
library(reshape2)
library(patchwork)
library(png)
library(grid)
library(gridBase)
library(gridExtra)
library(dplyr)
library(cowplot)
library(kableExtra)
library(xlsx)

library(ggpubr) #for arranging ggplots

@@

The approach will use the same glm models as in Chapter 2. A motivation for addressing these questions is that the variance of a quantitative trait can impact the evolution.

@@

# Q12 Variance Plots

@@

@@

#strain = fixed, 
varMLH1.M2.glm.strain.fixed.full <- glm(var ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

@@

**M4-M5 male**

```{r Male.specific.MLHvar.glm, echo=FALSE, eval=TRUE}

@@

**M2**

```{r cVMLH1.glm.M2, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")
#AP_mouse_table.HQ_target <- AP_mouse_table.HQ_target[ ! AP_mouse_table.HQ_target$strain %in% remove_strains, ]
#kkll <- DF.HetC.MixedModel.HQ[! DF.HetC.MixedModel.HQ$strain %in% remove_strains,]

#strain = fixed, 
cVMLH1.M2.glm.strain.fixed.full <- glm(cV ~ subsp * sex * strain, data= FULL_mouse_table_Q12.MLH1[  ! FULL_mouse_table_Q12.MLH1$strain %in% remove_strains, ]   )

@@

**M3**

```{r cVMLH1.glm.M3, echo=FALSE, eval=TRUE}
remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC")

@@

**M4-M5 female**

```{r Female.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

@@

**M4-M5 male**

```{r Male.specific.MLHcV.glm, echo=FALSE, eval=TRUE}

@@

sum_cVMLH1.male.glm.M4.2 <- summary(cVMLH1.male.glm.M4.2)

round(sum_cVMLH1.male.glm.M4.2$coefficients[,c(1,4)],5)

@@

I would suggest the following approach:
1. Consider only high-quality cells.

2. Pool cells within a strain and within a sex and compute one estimate of the variance in MLH1 count for each sex in each strain.

@@

4. For each permutation, compute the among-strain variance in the MLH1 count variance (separately for each sex).
Compute the among-strain variance in the MLH1 count variance (separately for each sex) in your original data.
Compare your observed variance with the null distribution generated by permutations to compute a p-value (one for each sex).

-->

@@

We'd like to know if the set of strains has bigger differences in between-cell variance than expected by chance. Your liner models provide one test, but they could suffer from small sample sizes per mouse. We can conduct a permutation test that enables you to pool all cells in each strain to increase the sample size.

#Permutation Step Up

Here's the permutation test I was imagining:
1. Separate the sexes.

2. Pool all cells within a strain and compute the **variance for each strain**.

3. Compute the **variance across strains** (VAS) of the variance within strains. This will leave you with a single estimate for the set of strains.

4. Randomly permute cells across strains. (pool all male cells, pool all female cells)

5. For the permuted data set, pool all cells within a strain and compute the variance for each strain. (compute a variance for a strain, var( sample( all.female.cells, N.WSB.f ) ) ).

6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.

7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.

8. Compare observed variance in variances to null distribution.

#observed values

@@

                        Nmice = length(unique(mouse)),
                        Ncells  = length(nMLH1.foci),
                        mean_co = as.numeric(format(round(mean(nMLH1.foci), 3), nsmall=3)),
                        cV = cv(nMLH1.foci),
                        var = as.numeric(format(round( var(nMLH1.foci),3), nsmall=3)),
                        sd   = round(sd(nMLH1.foci), 3),
                        se   = round(sd / sqrt(Ncells), 3)
)

@@

#remove extra strains
remove.strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
vals_Q12.MLH1 <- OG.vals_Q12.MLH1[!OG.vals_Q12.MLH1$strain %in% remove.strains, ]

#for males and females, calq the total variance across the strain-specific var
Obs.f.VAS <- var( ( vals_Q12.MLH1 %>% filter(sex == "female") )$var )
Obs.m.VAS <- var( ( vals_Q12.MLH1 %>%  filter(sex == "male") )$var )

@@

These are the female strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$var`, the variance across strains (VAS) is `r Obs.f.VAS`.

These are the male strain variances:  `r ( vals_Q12.MLH1 %>% filter(sex == "male") )$var`, the variance across strains (VAS) is `r Obs.m.VAS`. (For `r ( vals_Q12.MLH1 %>% filter(sex == "female") )$strain` respectively)

```{r str.of.DF, echo=FALSE, eval=TRUE}
#redundant
#remove_strains <- c("TOM","AST","CZECH","CAST","HMI","SPRET","CAROLI", "F1", "PERC", "SPIC", "MOLF")
#Q12.permute.DF <- Q12.MLH1_data[!Q12.MLH1_data$strain %in% remove_strains, ]

#Data structure (from the cell level data)
len.WSB.female <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "female") )$fileName )
len.WSB.male <- length( ( Q12.MLH1_data %>% filter(strain == "WSB") %>% filter(sex == "male") )$fileName )

@@

<!--
random permutations below, (COMPUTE the VAS)

For the permuted data set, **pool all cells within a strain (permuted strain ie defined by sample size?) and compute the variance for each strain**.

6. For the permuted data set, compute the variance across strains of the variance within strains. This will leave you with a single estimate for the set of strains.

7. Repeat 4-6 10,000 times to establish a null distribution for the variance in variances.

8. Compare observed variance in variances to null distribution.  -->

# Permutation Output

```{r permutations2, echo=FALSE}
#I made a DF of strain X sex specific variance

nrepeat = 10000

#I need to sample within strain? -- then I feel like I need to change the sample size -- otherwise I'm just sampling all the same episodes

permuted.var <- data.frame( "wsb_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[ (Q12.MLH1_data$sex == "female")], len.WSB.female, replace = FALSE) ) ),
                            "wsb_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.WSB.male, replace = FALSE) ) ), 
  "lew_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.LEW.female, replace = FALSE) ) ),

  "lew_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.LEW.male, replace = FALSE) ) ),

  "g_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.G.female, replace = FALSE) ) ), 
  "g_m.var" =  replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.G.male, replace = FALSE) ) ),

  "pwd_f.var" = replicate(n=nrepeat, var( sample( Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.PWD.female, replace = FALSE) ) ), 
  "pwd_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.PWD.male, replace = FALSE))),

  "kaz_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.KAZ.female,replace = FALSE) ) ),

  "kaz_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.KAZ.male, replace = FALSE)) ),

  "skive_f.var" = replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"],len.SKIVE.female,replace = FALSE) ) ),

  "skive_m.var" =  replicate(n=nrepeat,var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"],len.SKIVE.male, replace = FALSE) ) ),

  "msm_f.var" = replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "female"], len.MSM.female, replace = FALSE) ) ), "msm_m.var" =  replicate(n=nrepeat, var(sample(Q12.MLH1_data$nMLH1.foci[Q12.MLH1_data$sex == "male"], len.MSM.male, replace = FALSE) )   )  
  )

```{r functions.Chap3, echo=FALSE}
#formate the permuted DF, adding strain and sex to the DF

add_sex_chp3 <- function(dat){
  dframe <- dat
  dframe$sex <- "other"
  dframe$sex <- ifelse(grepl("f.", dframe$names), "female", 
                       ifelse(grepl("m.",dframe$names), "male", "other"))
  dframe$sex<- factor(dframe$sex, ordered = TRUE, levels =c( "female", "male"))
  return(dframe)
}

add_strain_Chap3 <- function(dat){
  dframe <- dat
  dframe$strain <- "other"
  dframe$strain <- ifelse(grepl("wsb", dframe$names), "WSB",
                    ifelse(grepl("g",dframe$names), "G",
                   ifelse(grepl("lew", dframe$names), "LEW",
                     ifelse(grepl("perc", dframe$names), "PERC",
                     ifelse(grepl("msm", dframe$names), "MSM",
                 ifelse(grepl("molf", dframe$names), "MOLF",
                     ifelse(grepl("skive", dframe$names), "SKIVE",      
                      ifelse(grepl("pwd", dframe$names), "PWD",
                  ifelse(grepl("kaz", dframe$names), "KAZ",
                   ifelse(grepl("tom", dframe$names), "TOM",
                  ifelse(grepl("ast", dframe$names), "AST",
                   ifelse(grepl("czech", dframe$names), "CZECH",
                    ifelse(grepl("cast", dframe$names), "CAST",
                   ifelse(grepl("hmi", dframe$names), "HMI",
                       ifelse(grepl("_SPRET_", dframe$names), "SPRET",
                       ifelse(grepl("_SPI_", dframe$names), "SPIC",
                                                                                                                                          ifelse(grepl("spic", dframe$names), "SPIC",
                                                                                                                                                 ifelse(grepl("_CAROLI_", dframe$names), "CAROLI",
                                 ifelse(grepl("F1_", dframe$names), "F1",         
                                   "other")))))))))))))))))))

  #the ordering factor below deletes all strain entries
  dframe$strain<- factor(dframe$strain, ordered = FALSE, levels =c( "WSB", "G", "LEW", "PERC",
                                                                   "PWD", "MSM", "MOLF","SKIVE", "KAZ", "TOM", "AST","CZECH",
                                                                   "CAST", "HMI",
                                                                   "SPRET", "SPIC", "CAROLI", "F1", "other") )
  return(dframe)
}

```{r pvals.VAS, echo=FALSE, warning=F, message=FALSE,  fig.width = 5, fig.height=5}

#need to merge all the pvales into a DF
#for(u in unique( OG.vals_Q12.MLH1$strain[ ( (OG.vals_Q12.MLH1$sex == "female") & (OG.vals_Q12.MLH1$species == "M.musculus") )] ) ){
#  print(u)
#  }

transposed.permuted <- t(permuted.var)

permuted.DF <- data.frame(names = row.names(transposed.permuted), transposed.permuted); rownames(transposed.permuted) <- NULL

permuted.DF <- add_sex_chp3(permuted.DF)
permuted.DF <- add_strain_Chap3(permuted.DF)

# 
VAS.F.permuted.DF <- permuted.DF %>% filter(sex == "female")
VAS.M.permuted.DF <- permuted.DF %>% filter(sex == "male")

#sampling.size <- nrepeat-3
#careful aboout matching the size of the dataframe!!

list.VAS_F <- c()
for(ff in 1:length(VAS.F.permuted.DF[,2:9997]) ){
  list.VAS_F[ff]  <- var( VAS.F.permuted.DF[,ff] )
}
#list of VAS
list.VAS_F.DF <- data.frame( "female.VAS" =  list.VAS_F)

list.VAS_M <- c()
for(mm in 1:length(VAS.M.permuted.DF[,2:9997]) ){

  perm.VAS.mm <- var( VAS.M.permuted.DF[,mm] )
  list.VAS_M[mm]  <- perm.VAS.mm

}
#p value is calculated for the distribution.

list.VAS_M.DF <- data.frame("male.VAS" = list.VAS_M)
#cbind the 2 VAS DF
Full.VAS <- cbind( list.VAS_F.DF, list.VAS_M.DF)

male.plot <- ggplot(Full.VAS, aes(x = male.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.m.VAS ) + ggtitle("Male Variance Across Species")

female.plot <- ggplot(Full.VAS, aes(x = female.VAS)) + geom_histogram() + geom_vline(xintercept = Obs.f.VAS ) + ggtitle("Female Variance Across Species")

male.plot
female.plot
```

The plots above show the distributions of permuted VAS for 10,000 permutations, the vertical line shows the observed VAS.

The proportion of permutations which are less than the oberserved VAS are `r sum(  Full.VAS$male.VAS <= Obs.m.VAS ) / nrepeat` and `r sum(  Full.VAS$female.VAS <= Obs.f.VAS ) / nrepeat`  for male and female respectively

```{r old.pval, echo=FALSE, eval=F}

F.strain.var <- c(0)
counter = 1
for(u in unique(permuted.DF$strain[permuted.DF$sex == "female"])){
  # print(u)

  F.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "female")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "female") )$var  ) / nrepeat
  counter = counter + 1
  #print(counter)
}
names(F.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "female"]  )

#male loop
M.strain.var <- c(0)
counter = 1
for(u in unique(permuted.DF$strain[permuted.DF$sex == "male"])){
  # print(u)
  # print( sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= (    OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat  )

  M.strain.var[counter] <-  sum( ( permuted.DF %>% filter(sex == "male")  %>% filter(strain == u )  )[,2:1000]  <= ( OG.vals_Q12.MLH1 %>% filter(strain == u) %>% filter(sex == "male") )$var  ) / nrepeat

  counter = counter + 1
  #print(counter)
}
names(M.strain.var) <- unique( permuted.DF$strain[permuted.DF$sex == "male"]  )

df.M.strain.var <- data.frame( vals = M.strain.var, strain = names(M.strain.var) )
df.F.strain.var <- data.frame( vals = F.strain.var, strain = names(F.strain.var) )

df.M.strain.var$sex <- "male"
df.F.strain.var$sex <- "female"

row.names(df.M.strain.var) <- c()
row.names(df.F.strain.var) <- c()

Strain.var.full <- rbind(df.M.strain.var, df.F.strain.var)

```{r p.vals, echo=FALSE, eval=FALSE}

p.val.WSB.f <- sum( (permuted.var$wsb.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var )) / nrepeat

p.val.WSB.m <- sum( (permuted.var$wsb.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )) / nrepeat

p.val.G.f <- sum( (permuted.var$g.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )) / nrepeat
p.val.G.m <- sum( (permuted.var$g.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )) / nrepeat

p.val.LEW.f <- sum( (permuted.var$lew.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var )) / nrepeat
p.val.LEW.m <- sum( (permuted.var$lew.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var )) / nrepeat

p.val.KAZ.f <- sum( (permuted.var$kaz.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )) / nrepeat
p.val.KAZ.m <- sum( (permuted.var$kaz.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )) / nrepeat

p.val.SKIVE.f <- sum( (permuted.var$skive.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )) / nrepeat
p.val.SKIVE.m <- sum( (permuted.var$skive.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )) / nrepeat

p.val.PWD.f <- sum( (permuted.var$pwd.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )) / nrepeat
p.val.PWD.m <- sum( (permuted.var$pwd.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )) / nrepeat

#calculate p values
p.val.MSM.f <- sum( (permuted.var$msm.f.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )) / nrepeat

p.val.MSM.m <- sum( (permuted.var$msm.m.var <= ( OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )) / nrepeat

```{r plots.var.dist, echo=FALSE, eval=F, warning=FALSE, message=FALSE}

#good ref
#http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/81-ggplot2-easy-way-to-mix-multiple-graphs-on-the-same-page/

wsb.fff <- ggplot(permuted.var, aes(x = wsb_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "female") )$var ) +  
  geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

wsb.mmm <- ggplot(permuted.var, aes(x = wsb_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "WSB") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "WSB") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

lew.fff <- ggplot(permuted.var, aes(x = lew_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "female") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#lew female observed var is greater than randomized
lew.mmm <- ggplot(permuted.var, aes(x = lew_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "LEW") %>% filter(sex == "male") )$var ) + geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "LEW") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

g.fff <- ggplot(permuted.var, aes(x = g_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#lew female observed var is greater than randomized
g.mmm <- ggplot(permuted.var, aes(x = g_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "G") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "G") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

kaz.fff <- ggplot(permuted.var, aes(x = kaz_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

kaz.mmm <- ggplot(permuted.var, aes(x = kaz_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "KAZ") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "KAZ") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

pwd.fff <- ggplot(permuted.var, aes(x = pwd_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

#pwd female oberserved variance is lower than the permutations.
pwd.mmm <- ggplot(permuted.var, aes(x = pwd_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "PWD") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "PWD") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

skive.fff <- ggplot(permuted.var, aes(x = skive_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

skive.mmm <- ggplot(permuted.var, aes(x = skive_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "SKIVE") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "SKIVE") ) , aes(x=25, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

msm.mmm <- ggplot(permuted.var, aes(x = msm_m.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "male") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "male") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

msm.fff <- ggplot(permuted.var, aes(x = msm_f.var))+ geom_histogram() + geom_vline(xintercept = (OG.vals_Q12.MLH1 %>% filter(strain == "MSM") %>% filter(sex == "female") )$var )+ geom_text(data= ( Strain.var.full %>% filter(sex == "female") %>% filter(strain == "MSM") ) , aes(x=18, y=80, label=vals),  colour="black", inherit.aes=FALSE, parse=FALSE)

@@

permutation.results <- ggarrange(wsb.fff, wsb.mmm,
          lew.fff, lew.mmm,
          g.fff, g.mmm, 
          kaz.fff, kaz.mmm,
          pwd.fff, pwd.mmm,
          skive.fff, skive.mmm, 
           msm.fff, msm.mmm + rremove("x.text"), 
          ncol = 4, nrow = 4)

annotate_figure(

  permutation.results, top = text_grob("permutations", color = "Black", face = "bold", size = 14) )
